### 2021-07-02 app_ch3.tex
got symbols working in the toc.

\documentclass[11pt]{article}
% General document formatting
\usepackage[margin=0.75in]{geometry}
\usepackage[parfill]{parskip}
\usepackage[utf8]{inputenc}
\usepackage{subfig}         % side-by-side figures 
% Related to math
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{titling}
\usepackage{hyperref}
\usepackage{wrapfig}
\usepackage{algpseudocode}
\usepackage{mathtools} % for colon equals

\usepackage{booktabs} % for wrapping tabulars in accord with
\bibliographystyle{agu}
\setlength{\droptitle}{-5em}   % This is your set screw

%\usepackage[math]{kurier}
\newcommand\be{\begin{equation}} % shortcut to start eq envs 
\newcommand\ee{\end{equation}}   % shortcut to end eq envs
\newcommand\ol{\overline}        % shortcut to draw overline 
\newcommand\bra{\langle}
\newcommand\ket{\rangle}
\newcommand\El{\mathcal{L}}
\newcommand\tg{\tilde{g}}
\newcommand\tG{\tilde{G}}
\date{} % make the date field blank

% for pseudo code -- example from https://tex.stackexchange.com/questions/163768/write-pseudo-code-in-latex
\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

\begin{document}
\title{Supplementary Material for ``Joint stochastic bedload transport and bed elevation model: variance regulation and power-law rests" by J. Kevin Pierce and Marwan A. Hassan}
\maketitle

@@

\bibliography{biblio}
\end{document}

>>>>

%%!TEX root = diss.tex

\chapter{Ch. 3 calculations}

@@

### 2021-07-02 app_ch5.tex
got symbols working in the toc.

\chapter{Appendices}
\section{Calculations of chapter 1:}
\section{Calculations of chapter 2: From particle dynamics to the sediment flux}
\section{Calculations of chapter 3:}

\section{Calculations of Chapter 4: }

\section{Strategies for Monte Carlo simulation}

\subsection{Derivation of Master Equation}
To derive the master equation from \ref{eq:langevin}, we temporarily consider the Gaussian white noise (GWN) $\xi(t)$ as a Poisson jump process having rate $r$ and jumps $ \sqrt{2 d} h$ with $h$ distributed as $f(h)$. We will later take a GWN limit on this noise. With this assumption, integrating \ref{eq:langevin} over a small time interval $\delta t$ considering the Ito interpretation for the collision term provides
\be     
u(t+\delta t) =
\begin{cases}
	u(t) + \gamma \delta t & \text{with probability } 1- r \delta t - \nu \delta t\\
	u(t) + \sqrt{2 d} h &  \text{with probability } r \delta t \\
	\ve u(t) &  \text{with probability } \nu \delta t
\end{cases}
\ee

Considering the probability $P(u,t+\delta)$ as sum over possible paths from $P(u,t)$ develops 
\begin{align} P(u,t+\delta t) =
	(1- r \delta t - \nu \delta t)\int_{-\infty}^\infty dw  \delta(u-w-\gamma \delta t) P(w,t)\\ 
	+  r \delta t \int_{-\infty}^\infty dw \int_{-\infty}^\infty dh f(h) \delta (u - w - \sqrt{2 d}h) P(w,t) \\ 
	+ \nu \delta t \int_{-\infty}^\infty dw \int_0^1 d\ve \rho(\ve)  \delta(u-w\ve) P(w,t).
\end{align}
Evaluating all integrals over $\delta$-functions provides 
\begin{align} P(u,t+\delta t) = (1-r \delta t - \nu \delta t)P(u-\gamma \delta t, t) \\+ r \delta t\int_{-\infty}^\infty dh f(h) P(u+\sqrt{2d}h) \\+ \nu \delta t \int_0^1 \frac{d\ve}{\ve} \rho(\ve) P\big(\frac{u}{\ve}\big).\end{align}
Finally, we take $\delta t \rightarrow 0$ and limit the Poisson noise involving $\sqrt{2d}$ to a Gaussian white noise by taking $r \rightarrow \infty$ as $h \rightarrow 0$ such that $h^2 r = 1$ \cite{VanKampen1983}. This process finally obtains the master equation (\ref{eq:master}).
\subsection{Derivation of Steady-state solution}

@@

\subsection{Calculation of the moments}
Taking (\ref{eq:master}), multiplying by $u^k$, integrating over all space, and taking account of normalization of $P(u)$ provides a recursion relation for the moments: 
\be 0 = Dk (k-1)\langle u^{k-2} \rangle + \Gamma k \langle u^{k-1} \rangle + \nu (\ve^k-1)\langle u^k \rangle. \ee
$k=1$ provides the mean
\be \langle u \rangle = \frac{\Gamma}{\nu(1-\ve)} = \frac{\gamma}{1-\ve}\ee
while $k=2$ provides the second moment
\be \langle u^2 \rangle = 2 \frac{d + \gamma \langle u \rangle}{1-\ve^2}, \ee
leading to the velocity variance
\be \sigma_u = \sqrt{\frac{2 d + \gamma^2}{1-\ve^2}}.\ee
\subsection{Weak and strong collision limits}
Now we demonstrate that weak collisions imply a Gaussian-like distribution for sediment velocities. The limit is challenging since the steady-state distribution \ref{eq:steadystate} and the moments above all diverge as $\ve \rightarrow 1$.
Following \cite{Hall1970}, this suggests normalizing the distribution $P(u)$ using 
\be z = \frac{u - \bar{u}}{\sigma_u} \ee
and 
\be Q(z) = \sigma_u P(u)\ee
to seek a differential equation for $Q(z)$ with manageable behavior as $\ve \rightarrow 1$.
Incorporating this transformation into (\ref{eq:master}) provides a ``normalized" Master equation
\be (1-\ve^2) \frac{d}{2d + \gamma} Q''(z) - \frac{\gamma\sqrt{1-\ve^2}}{\sqrt{2d+\gamma^2}}Q'(z) - Q(z) + \frac{1}{\ve} Q\Big(z + \Big[\frac{1-\ve}{\ve}z + \frac{\gamma\sqrt{1-\ve^2}}{\ve\sqrt{2d+\gamma^2}} \Big]\Big) = 0 \ee
This equation remains exact and is only a change of variables from (\ref{eq:master}).
Now we approximate the equation for $\ve \rightarrow 1$ by expanding the final term to second order around $z=0$ before setting $\ve = 1$, obtaining
\be Q''(z) + z Q'(z) + Q(z) = 0, \ee
which is the classic Ornstein-Uhlenbeck Fokker-Planck equation whose solution is the standard normal distribution for $Q(z)$. This solution provides \ref{eq:gaussian} when transformed back to the original variables $P(u)$ and $u$.

>>>>

\chapter{Ch. 5 calculations}

\section{Derivation of Master Equation}
To derive the master equation from \ref{eq:langevin}, we temporarily consider the Gaussian white noise (GWN) $\xi(t)$ as a Poisson jump process having rate $r$ and jumps $ \sqrt{2 d} h$ with $h$ distributed as $f(h)$. We will later take a GWN limit on this noise. With this assumption, integrating \ref{eq:langevin} over a small time interval $\delta t$ considering the Ito interpretation for the collision term provides
\be     
u(t+\delta t) =
\begin{cases}
	u(t) + \gamma \delta t & \text{with probability } 1- r \delta t - \nu \delta t\\
	u(t) + \sqrt{2 d} h &  \text{with probability } r \delta t \\
	\ve u(t) &  \text{with probability } \nu \delta t
\end{cases}
\ee

Considering the probability $P(u,t+\delta)$ as sum over possible paths from $P(u,t)$ develops 
\begin{align} P(u,t+\delta t) =
	(1- r \delta t - \nu \delta t)\int_{-\infty}^\infty dw  \delta(u-w-\gamma \delta t) P(w,t)\\ 
	+  r \delta t \int_{-\infty}^\infty dw \int_{-\infty}^\infty dh f(h) \delta (u - w - \sqrt{2 d}h) P(w,t) \\ 
	+ \nu \delta t \int_{-\infty}^\infty dw \int_0^1 d\ve \rho(\ve)  \delta(u-w\ve) P(w,t).
\end{align}
Evaluating all integrals over $\delta$-functions provides 
\begin{align} P(u,t+\delta t) = (1-r \delta t - \nu \delta t)P(u-\gamma \delta t, t) \\+ r \delta t\int_{-\infty}^\infty dh f(h) P(u+\sqrt{2d}h) \\+ \nu \delta t \int_0^1 \frac{d\ve}{\ve} \rho(\ve) P\big(\frac{u}{\ve}\big).\end{align}
Finally, we take $\delta t \rightarrow 0$ and limit the Poisson noise involving $\sqrt{2d}$ to a Gaussian white noise by taking $r \rightarrow \infty$ as $h \rightarrow 0$ such that $h^2 r = 1$ \cite{VanKampen1983}. This process finally obtains the master equation (\ref{eq:master}).
\section{Derivation of Steady-state solution}

@@

\section{Calculation of the moments}
Taking (\ref{eq:master}), multiplying by $u^k$, integrating over all space, and taking account of normalization of $P(u)$ provides a recursion relation for the moments: 
\be 0 = Dk (k-1)\langle u^{k-2} \rangle + \Gamma k \langle u^{k-1} \rangle + \nu (\ve^k-1)\langle u^k \rangle. \ee
$k=1$ provides the mean
\be \langle u \rangle = \frac{\Gamma}{\nu(1-\ve)} = \frac{\gamma}{1-\ve}\ee
while $k=2$ provides the second moment
\be \langle u^2 \rangle = 2 \frac{d + \gamma \langle u \rangle}{1-\ve^2}, \ee
leading to the velocity variance
\be \sigma_u = \sqrt{\frac{2 d + \gamma^2}{1-\ve^2}}.\ee
\section{Weak and strong collision limits}
Now we demonstrate that weak collisions imply a Gaussian-like distribution for sediment velocities. The limit is challenging since the steady-state distribution \ref{eq:steadystate} and the moments above all diverge as $\ve \rightarrow 1$.
Following \cite{Hall1970}, this suggests normalizing the distribution $P(u)$ using 
\be z = \frac{u - \bar{u}}{\sigma_u} \ee
and 
\be Q(z) = \sigma_u P(u)\ee
to seek a differential equation for $Q(z)$ with manageable behavior as $\ve \rightarrow 1$.
Incorporating this transformation into (\ref{eq:master}) provides a ``normalized" Master equation
\be (1-\ve^2) \frac{d}{2d + \gamma} Q''(z) - \frac{\gamma\sqrt{1-\ve^2}}{\sqrt{2d+\gamma^2}}Q'(z) - Q(z) + \frac{1}{\ve} Q\Big(z + \Big[\frac{1-\ve}{\ve}z + \frac{\gamma\sqrt{1-\ve^2}}{\ve\sqrt{2d+\gamma^2}} \Big]\Big) = 0 \ee
This equation remains exact and is only a change of variables from (\ref{eq:master}).
Now we approximate the equation for $\ve \rightarrow 1$ by expanding the final term to second order around $z=0$ before setting $\ve = 1$, obtaining
\be Q''(z) + z Q'(z) + Q(z) = 0, \ee
which is the classic Ornstein-Uhlenbeck Fokker-Planck equation whose solution is the standard normal distribution for $Q(z)$. This solution provides \ref{eq:gaussian} when transformed back to the original variables $P(u)$ and $u$.

### 2021-07-02 ch6conc.tex
got symbols working in the toc.

\chapter{Summary and future work}
\label{ch:conc}

@@

\subsection{Probability distribution of the sediment flux}

@@

\subsection{Quantification of the control of bed elevation fluctuations on sediment transport fluctuations}

\subsection{Understanding of how sediment burial affects the downstream spreading of sediment tracer particles}

@@

\section{Closure: Complexity vs Realism}

>>>>

\chapter{Summary and future work}
\label{ch:conc}
\section{The overall strategy}
\subsection{Langevin and master equations}
\subsection{Idealized noises and their combinations}

@@

\subsection{Probability distribution of the sediment flux from micromechanics of particle trajectories}

@@

\subsection{Quantification of the control of bed elevation fluctuations over sediment transport fluctuations}

\subsection{Predicting how sediment burial affects the downstream spreading of sediment tracer particles}

@@

\section{Conclusion and next steps}

### 2021-07-02 pref.tex
got symbols working in the toc.

\chapter{Preface: complementary traditions}

This thesis deals with the transport of coarse sediment in flowing water from the perspective that sediment transport results from the movements of individual grains.
The research has been motivated by the belief that geomorphology as a science will benefit from mechanistic, process-based, mathematical models of sediment transport processes.

@@

Given the complexity of geomorphology problems, a complete reduction of geomorphology to physics would be narrow-minded, but we can nonetheless borrow ways of thinking.
An extremely successful approach in physics has been to construct idealized models with little intention of direct realism, study them deeply to obtain complete understanding, and then, decades or centuries later, to embellish these models with more sophisticated features to describe real-world phenomena.
An example is the block on the spring, the simple harmonic oscillator, which starts as a toy model studied by every first year physics student, but somehow shows up in the deepest inquiries of theoretical physics, from quantum matter to cosmological inflation.
This thesis applies the simple harmonic oscillator approach to some problems in Earth science.
The objective is to build up simple archetypes which can be embellished later, not to build \textbf{the model} of fluvial geomorphology. 
That's for later. (And it's not for me!) CHEERS.

>>>>

\chapter{Preface}

This thesis is entirely the original research of Kevin Pierce, including all figures, writing, and calculations.
The supervisory committee and especially the research supervisor Marwan Hassan provided guidance in the research.
The thesis includes two published works, Chapters \ref{ch:bedEle} and \ref{ch:downdiff}, available in \textit{Journal of Geophysical Research: Earth Surface} and \citet{Geophysical Research Letters} respectively, with Marwan Hassan as coauthor \citep{Pierce2020a,Pierce2020b}. Chapters \ref{ch:flux} and \ref{ch:langevin} will be submitted for publication subsequent to the thesis submission.

This thesis deals with the transport of coarse sediment in flowing water from the perspective it fundamentally results from the movements of individual grains.
The research has been motivated by the belief that geomorphology as a science will benefit from mechanistic, process-based, mathematical models of sediment transport processes.

@@

Given the complexity of geomorphology problems, a complete reduction of geomorphology to physics would be narrow-minded, but we can nonetheless borrow ways of thinking.
An extremely successful approach in physics has been to construct idealized models with little intention of direct realism, study them deeply to obtain complete understanding, and then, decades or centuries later, to embellish these models with more sophisticated features to describe real-world phenomena.
An example is the block on the spring, the simple harmonic oscillator, which starts as a toy model studied by every first year physics student, but somehow shows up in the deepest inquiries of theoretical physics, from quantum matter to cosmological inflation.
This thesis applies this ``simple harmonic oscillator approach" to some problems in Earth science.
The objective is to build up simple archetypes which can be embellished later, not to build \textbf{the model} of fluvial geomorphology. 
That's for later! (And not for me.) CHEERS.
